
\newpage
\begin{thebibliography}{4}

\bibitem{BrownSchnetz} Brown, F., Schnetz, O.: Modular forms in quantum field theory. Communications in Number Theory and Physics 7(2), 293--325 (2013)

\bibitem{Colton1} Colton, S.: Automated Theory Formation in Pure Mathematics. 
PhD. Thesis, Department of Artificial Intelligence, University of Edinburgh (2001)

\bibitem{Colton2} Colton, S.: Automated Conjecture Making in Number Theory using HR, Otter and Maple. Journal of Symbolic Computation 39(5), 593--615 (2005) 

\bibitem{Colton3} Colton, S.: Refactorable Numbers - A Machine Invention. Journal of Integer Sequences 2, 99.1.2 (1999)

\bibitem{CostaTschinkel} Costa, E., Tschinkel, Y.: Variation of Neron-Severi ranks of reductions of K3 surfaces. Experimental Mathematics 23, pp. 475--481  (2014)

\bibitem{DurfeeMartin} Durfee, C., Martin, K.: Distinguishing graphs with zeta functions and generalized spectra. Linear Algebra Appl. 481, pp. 54--82 (2015)

\bibitem{EoM} Encyclopedia of Mathematics, \url{http://www.encyclopediaofmath.org/}

\bibitem{GaitsgoryLurie} Gaitsgory, D., Lurie, J.: Weil's Conjecture for Function Fields. Preprint available at \url{http://www.math.harvard.edu/~lurie/papers/tamagawa.pdf}


\bibitem{Ganesalingam} Ganesalingam, M.: The Language of Mathematics - A Linguistic and Philosophical Investigation. Theoretical Computer Science and General Issues, Vol. 7805. Springer-Verlag Berlin Heidelberg (2013)


\bibitem{GranvilleSound} Granville, A., Soundararajan, K.: Pretentious multiplicative functions and an inequality for the zeta-function. In: De Koninck, J., Granville, A., Luca, F. (eds.): Anatomy of Integers, CRM Proceedings and Lecture Notes, Vol. 46 (2008)

\bibitem{Harvey1} Harvey, D.: Counting points on hyperelliptic curves in average polynomial time. Ann. of Math. (2) 179(2), 783--803 (2014)

\bibitem{Harvey2} Harvey, D.: Computing zeta functions of arithmetic schemes.
Proc. Lond. Math. Soc. 111, no. 6, 1379--1401 (2015)


\bibitem{Kedlaya} Kedlaya, K.S.: Computing Zeta Functions via p-adic Cohomology. In: Buell (ed.); Algorithmic Number Theory. LNCS, vol. 3076, pp. 1--17. Springer Berlin Heidelberg (2004) 

\bibitem{LarsonVC} Larson, C. E., Van Cleemput, N.: Automated Conjecturing I: Fajtlowicz's Dalmatian Heuristic Revisited. Artificial Intelligence 231, 17--38 (2016)

\bibitem{LMFDB} The {LMFDB Collaboration}: The L-functions and Modular Forms Database. \url{http://www.lmfdb.org}

\bibitem{Marcolli} Marcolli, M.: Feynman motives. World Scientific, Hackensack, NJ (2010)

\bibitem{OEIS} Online Encyclopedia of Integer Sequences, \url{http://oeis.org/}

\bibitem{Ronan} Ronan, M.: Symmetry and the Monster: One of the greatest quests of mathematics. Oxford University Press, UK (2006)

\bibitem{duSautoyICM} Add details here.

\bibitem{Toth} See \url{http://ttk.pte.hu/matek/ltoth/Toth_Pillai2_1996.pdf}

\bibitem{VollIntro} Voll, C.: An newcomer's guide to zeta functions of groups and rings. Preprint available at \url{https://arxiv.org/abs/0906.1832}

\bibitem{Whitcher} Whitcher, U." Mirror symmetry and K3 surface zeta functions. Presentation given at ICERM, Oct 2015. Slides available at \url{https://icerm.brown.edu/sp-f15-w2/}


\bibitem{Zeilberger} Zeilberger, D.: A holonomic systems approach to special functions identities. J. Comput. Appl. Math. 32(3), 321--368 (1990)



%%%%

%\bibitem{jour} Smith, T.F., Waterman, M.S.: Identification of Common Molecular Subsequences. J. Mol. Biol. 147, 195--197 (1981)

%\bibitem{lncschap} May, P., Ehrlich, H.C., Steinke, T.: ZIB Structure Prediction Pipeline: Composing a Complex Biological Workflow through Web Services. In: Nagel, W.E., Walter, W.V., Lehner, W. (eds.) Euro-Par 2006. LNCS, vol. 4128, pp. 1148--1158. Springer, Heidelberg (2006)

%\bibitem{book} Foster, I., Kesselman, C.: The Grid: Blueprint for a New Computing Infrastructure. Morgan Kaufmann, San Francisco (1999)

%\bibitem{proceeding1} Czajkowski, K., Fitzgerald, S., Foster, I., Kesselman, C.: Grid Information Services for Distributed Resource Sharing. In: 10th IEEE International Symposium on High Performance Distributed Computing, pp. 181--184. IEEE Press, New York (2001)

%\bibitem{proceeding2} Foster, I., Kesselman, C., Nick, J., Tuecke, S.: The Physiology of the Grid: an Open Grid Services Architecture for Distributed Systems Integration. Technical report, Global Grid Forum (2002)

%\bibitem{url} National Center for Biotechnology Information, \url{http://www.ncbi.nlm.nih.gov}

\end{thebibliography}
